---
output:
  pdf_document: default
  html_document: default
---
# Data transformation


## Brief View of the Dataset

Here is a brief view of the first 20 rows of the dataset.
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(DT)
vgsales = read_csv("vgsales-12-4-2019.csv")
datatable(head(vgsales,20),rownames = FALSE,options = list(scrollX = TRUE))
```


Let us also take a view at all variables of the data.
```{r}
colnames(vgsales)
```

Obviously, there are many redundant variables in our dataset. We will not be interested in investigating the url of the game, or the update information about the game, etc. Thus, we deleted a series of useless variables that include basename, ESRB_Rating, Developer, VGChartz_Score, Total_Shipped, Last_Update, url, status, Vgchartzscore, and img_url. This is what our updated dataset looks like:

```{r}
dataset = vgsales[,c("Rank","Name","Genre","Platform","Publisher","Critic_Score","User_Score","Global_Sales","NA_Sales","PAL_Sales","JP_Sales","Other_Sales","Year")]
head(dataset)
```

Note that all missing values will be treated in the next section. 