# Interactive component

```{r echo=FALSE}
datana=data.frame(genre=genre, sales = nasales)
datana=datana[order(-datana$sales),]
datajapan=data.frame(genre=genre, sales = jasales)
datajapan=datajapan[order(-datajapan$sales),]
dataeu=data.frame(genre=genre, sales = eusales)
dataeu=dataeu[order(-dataeu$sales),]
dataother=data.frame(genre=genre, sales = othersales)
dataother=dataother[order(-dataother$sales),]
```

```{r echo=FALSE, results="asis"}
library("jsonlite")
cat(
  paste(
  '<script>
    var datana = ',toJSON(datana),';
    var datajapan = ',toJSON(datajapan),';
    var dataeu = ',toJSON(dataeu),';
    var dataother = ',toJSON(dataother),';
  </script>'
  , sep="")
)
```

<script src="https://d3js.org/d3.v6.js"></script>



<button onclick="update(datana)">North America Sales</button>
<button onclick="update(dataeu)">Europe Sales</button>
<button onclick="update(dataother)">Other Sales</button>
<button onclick="update(datajapan)">Japan Sales</button>

<div id="my_dataviz"></div>
<script>
var margin = {top: 40, right: 30, bottom: 70, left: 60},
    width = 500 - margin.left - margin.right,
    height = 450 - margin.top - margin.bottom;


var svg = d3.select("#my_dataviz")
  .append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)
  .append("g")
    .attr("transform",
          "translate(" + margin.left + "," + margin.top + ")");

// X axis
var x = d3.scaleBand()
  .range([ 0, width ])
  .padding(0.2);
var xAxis = svg.append("g")
  .attr("transform", "translate(0," + height + ")")
;
  
// Y axis
var y = d3.scaleLinear()
  .range([ height, 0]);
var yAxis = svg.append("g")
  .attr("class", "myYaxis")


// A function that create / update the plot for a given variable:
function update(data) {

  x.domain(data.map(function(d) { return d.genre; }))
  xAxis.call(d3.axisBottom(x))
  .selectAll("text")
    .attr("y", 0)
    .attr("x", 9)
    .attr("dy", ".35em")
    .attr("transform", "rotate(90)")
    .style("text-anchor", "start");

  svg.append("text")             
      .attr("transform",
            "translate(" + (width/2) + " ," + 
                           (height + margin.top + 20) + ")")
      .style("text-anchor", "middle")
      .text("Genres");  
  
  y.domain([0, d3.max(data, function(d) { return d.sales }) ]);
  yAxis.transition().duration(1000).call(d3.axisLeft(y));

   svg.append("text")
      .attr("transform", "rotate(-90)")
      .attr("y", 0 - margin.left)
      .attr("x",0 - (height / 2))
      .attr("dy", "1em")
      .style("text-anchor", "middle")
      .text("Sales in Million");    



  
  var u = svg.selectAll("rect")
    .data(data)





  u
    .enter()
    .append("rect")
    .merge(u)
    .transition()
    .duration(1000)
      .attr("x", function(d) { return x(d.genre); })
      .attr("y", function(d) { return y(d.sales); })
      .attr("width", x.bandwidth())
      .attr("height", function(d) { return height - y(d.sales); })
      .attr("fill", "steelblue")


  u.
  on("mouseover", function(d){d3.select(this).style('fill','red');
})
  .on("mouseout", function(d){d3.select(this).style("fill","steelblue");
});
    

}


update(datajapan)

</script>





\
By clicking the buttom, we can see the sales of genres with decreasing order for each region. In the D3 graph, we can clearly see the the sales trend changes in an interactive way. We could eaisly find the most popular and the least popular games for different regions and compare the sales differences. In particular, the top 3 genres for all regions are Action, Shooter, Role-playing. Additionally, Action is the most popular genre all over the world. For new video game producers, they could target on Action game to reach high level of sales. 
